\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsfonts}
\title{CSCI 510, Fall 2016, Homework \# 4}
\author{SOLUTION SKETCHES}
\date{Due date: Wednesday, November 2, Midnight} 

\begin{document}
\maketitle
\begin{enumerate}
\item Let $A$ be a Turing-recognizable language consisting of Turing
  machines, $\{\langle M_1 \rangle, \langle M_2\rangle, \ldots\}$,
  where every $M_i$ is a decider.  Prove that some decidable
  language $D$ is not decided by any decider $M_i$ whose description
  appears in $A$.

  \newcommand{\bd}{\begin{description}}
  \newcommand{\ed}{\end{description}}

  Since $A$ is Turing-recognizable, it is also Turing-enumerable, and
  so we assume that there is a machine to enumerate the machines
  $M_i$. 
  Description of machine $D$:
  
  \hrulefill
  \bd
  \item On input $w$:
    \bd
  \item Let $\ell=|w|$
  \item Run the enumerator for $A$ $\ell$ times to find machine $M_\ell$.
    \item Run machine $M_\ell$ on $w$.
  (This must halt, since $M_\ell$ is a decider.)
  \item If $M_\ell$ accepts, {\em reject}, else {\em accept}.
    \ed
    \ed
  \hrulefill

Machine $D$ is a decider since it always halts.  Machine $D$ cannot be
in the list that $A$ enumerates, since for each machine $M_i\in A$ it gives
different judgements on all strings of length $i$.

  

\item Let $A= \{\langle R\rangle | $ $R$ is a regular
  expression describing a language containing at least one string
  $w$ that has  $111$  as a substring ({\em i.e.} $w=x111y$ for
  some $x$ and $y$ $\}$.  Show that $A$ is decidable.

  We can turn any regular expression into a deterministic finite
  automaton.  Now the problem has been reduced to finding whether
  there exists a path from the start state of the DFA to an accept
  state with 111 along the path somewhere.

  This can be solved easily with a graph search algorithm that starts
  at the root and expands paths until all reach accept states, or all
  start looping (if the length of the loop is less than 3, loop one or
  two more times).  Examine all these paths to see if 111 shows up.


\item Let $A= \{\langle R, S\rangle |$ $R$ and $S$ are regular
  expressions and $L(R) \subseteq L(S)\}$.  Show that $A$ is
  decidable.

  Find DFAs for $L(R)$ and $L(S)$, and then find the DFA for the
  difference, $L(R)-L(S) = L(R)\cap \overline{L(S)}$.  There are
  terminating algorithms for each of these steps.  If this language is
  empty (which is decidable, see text), then $L(R)\subseteq L(S)$.
  
\end{enumerate}
\end{document}
