\documentclass{article}
\usepackage[margin=0.75in]{geometry}
\usepackage{alltt}
\usepackage{multicol}
\usepackage{hyperref}
\begin{document}

\centerline{\large \bf Syllabus}

\centerline{\bf CSCI 510, Automata and Formal Language Theory, Fall 2016}

\begin{itemize}

\item
{\bf Instructor:} Geoffrey Matthews, x3797, {\tt  geoffrey dot matthews at wwu dot edu}
\item
{\bf Office hours:} MTWF 10:00, CF 469
\item
{\bf Text:} 
{\em  Introduction to the Theory of Computation}, Michael Sipser, any edition.
  Note that the 3rd edition in paperback is only \$39 from Amazon.

\item {\bf Webpage:}  \url{wwu.instructure.com}:  homework assignments, grades
\item {\bf Repository:} \url{https://github.com/geofmatthews/csci510}:  handouts, lectures, code
\item
{\bf Lectures:} 
CF224, MTWF 9:00

\item
  {\bf Content:} We will cover regular languages, context free languages,
  and recursive languages by examining the properties of the languages
  and also the types of machines that recognize them.  Such machines
  include finite state machines, pushdown automata and Turing
  Machines.  We will also look at undecidable problems, the theory of
  NP-completeness, and discuss other meaningful ways to classify
  problems. 

\item {\bf Exams:} One midterm and one final.  You may bring two
  double-sided pages of notes to use during the exams.


\item {\bf Homework:} Homework assignments will be passed out
  regularly through the quarter.  Homework will be due at midnight on
  the due date.  Late work is accepted at a penalty of 25\% per each
  fraction of 24 hours late.  There may be an assignment due during
  dead week.


\item {\bf Grading:} \\
$0\leq F < 60 \leq D < 70 \leq C < 80 \leq B < 90 \leq A$\hfill
\begin{tabular}{|l|l|l|}\hline
Homework  & Midterm & Final\\\hline
50\% & 20\% & 30\%\\\hline
\end{tabular}

\item {\bf Academic dishonesty:} Academic dishonesty policy and
  procedure is discussed in the University Catalog, Appendix D.  All
  students should read this section of the catalog.  Academic
  dishonesty consists of misrepresentation by deception or other
  fraudulent means.

\item {\bf Collaboration:} Collaboration with your fellow students is
  a good way to learn.  Feel free to share ideas, solve problems, and
  discuss your programs with other students.  However, collaboration
  is {\em not} copying.

\item
{\bf Approximate Schedule:} The following schedule may be adjusted
radically depending on interests and problems as they occur.  


\begin{verbatim}
   September 2016     
Su Mo Tu We Th Fr Sa  
18 19 20 21 22 23 24  Chapters 0, Mathematical concepts
    October 2016      
25 26 27 28 29 30  1  Chapters 1, Regular languages
 2  3  4  5  6  7  8  Chapters 2, Context-free languages
 9 10 11 12 13 14 15  Chapters 3, Church-Turing Thesis
16 17 18 19 20 21 22  Chapters 4, Decidability
23 24 25 26 27 28 29  Review and Catchup, Midterm Friday October 28
   November 2016      
30 31  1  2  3  4  5  Chapters 5, Reducibility
 6  7  8  9 10 11 12  Chapters 6, Advanced topics
13 14 15 16 17 18 19  Chapters 7, Time complexity
20 21 22 23 24 25 26  Chapters 8, Space complexity
   December 2016      
27 28 29 30  1  2  3  Review and Catchup
 4  5  6  7  8  9 10  Final Monday December 5 8:00am
\end{verbatim}
\end{itemize}

\end{document}
